---
title: "Week 4 Assingment"
author: "Kellen Sorauf"
date: "February 4, 2020"
output:
  html_document: default
  pdf_document: default
---
# **This is R Markdown**
#### You can create nice publishable documents using R Markdown
You should be comfortable using R Markdown for your R assignments.  
You can find cheatsheets for R Markdown here
https://rmarkdown.rstudio.com/lesson-1.html

### Let's begin the assignment.
Import the data.table and ggplot2 packages.  If they are not already installed install: 
install.packages('ggplot2') install.packages('data.table')

```{r setup, warning=FALSE, echo=TRUE}
library(data.table)
library(ggplot2)
```

Import the Binomial.csv data. There are many ways import a .csv file, but fread() will import it as a data.table

```{r import data, echo = TRUE}
bi <- fread('C:/Users/ksorauf/OneDrive - Regis University/MSDS 600/19S8W1/Week 4/Data/Binomial.csv')
```

Calculate summary statistics and standard deviation. Plot the data as a scatter plot, histogram, and boxplot

```{r Binomial summary statistics and plots, echo=TRUE}

summary(bi)

sd(bi$V1)

plot(bi$V1, xlab = 'bi Index', ylab = 'bi values', main = 'Scatter plot of the Binomial dataset')

hist(bi$V1, xlab = 'bi values', ylab = 'Frequency of bi values', main = 'Histogram of the Binomial dataset')

boxplot(bi, xlab = 'bi', ylab = 'bi values', main = 'Boxplot of the Binomial dataset')
```

Load the six datasets in the week 4 materials.  Compute a set of descriptive statistics for each; including mean, standard deviation, minimum, maximum, a histogram plot, and any other descriptive statistic you might find meaningful - include these in your write-up.

Import the Ln.csv data.
Calculate summary statistics and standard deviation. Plot the data as a scatter plot, histogram, and boxplot.

```{r LN summary statistics and plots, echo=FALSE}

```

Import the BN1.csv and BN2.csv data and calculate summary statistics, standard deviation and plot the histograms individually. 
Plot the BN1 and BN2 together on one graph as a scatter plot, and as boxplots.

```{r BN summary statistics and plots, echo=FALSE}

```

Import the N1.csv and N2.csv data and calculate summary statistics, standard deviation and plot the histograms individually. 
Plot the BN1 and BN2 together on one graph as a scatter plot, and as boxplots.

```{r N summary statistics and plots, echo=FALSE}

```

What, if anything, can be said about the differences between N1 & N2?

Similarly, what can be said about the differences between BN1 & BN2?

What are the differences in the distributions, how do the outliers differ, what are the differences in the means, range, etc?

When considering the vector individually, check if the distribution is normal or not -- which stats would you use to determine this? 